---
title: "MSpallya Report"
author: "saurabh"
date: "30/06/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Libraries used
```{r collapse=TRUE, warning=FALSE,message=FALSE}
library(ggplot2)
library(pheatmap)
library(gplots)
require(made4)
require(heatmaply)
require(dplyr)
library(tidyr)
require(RColorBrewer)
library(DESeq2)
library(DESeq)
library(reshape)
library(tidyverse)
```


## This Markdown report is created for MSPALLYA 20 mosquito samples

```{r dff_merge_all}
dff_merge_all <- read.csv("dff_merge_all_removed.csv",sep="\t",header = TRUE)
data <- dff_merge_all
data$taxon_name <- NULL
data <- data.frame(data)
data <- data %>% drop_na()
```

## What is data we are using
```{r}
head(data)
```

## General Heatmap
```{r}
pheatmap(log2(data[1:150,1:20]),fontsize = 5, scale = "none", cluster_rows = F,cutree_rows = 5,cellwidth = 15,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = F, show_colnames = T, main = "1-150:Log2",fontsize_row = 10,ffontsize_col = 8)
```



## Data Summary
```{r}
summary(data)
```

## Setting data for DESeq analysis

```{r}
group = c(rep('G0',4),rep('G1',4),rep('G2',4),rep('G3',4),rep('G4',4))
colData <- read.csv("colData.csv",header = TRUE,sep="\t")
dds <- DESeqDataSetFromMatrix(countData = data, colData = colData,design=~stage)
```

## DESeq Analysis
```{r, warning=FALSE,message=FALSE}
dds <- DESeq(dds)
res <- results(dds, tidy=TRUE)
res <- as_tibble(res)
```

## summary and coldata printing
```{r}
head(counts(dds),n = 20)
colData(dds)
```


## DESeq Filtering based on padj < 0.05 
head and tail printed
```{r}
padj_05 <- res %>% filter(padj<0.05)
head(padj_05)
tail(padj_05)
```

# We can plot graphs for some bacteria

## Plotting Asasia with the values in each sample
```{r}
plotCounts(dds, gene="Asaia", intgroup="stage")
```

## Plotting Asaia using ( Box Plot)
```{r}
plotCounts(dds, gene="Asaia", intgroup="stage", returnData=TRUE) %>% 
  ggplot(aes(stage, count)) + geom_boxplot(aes(fill=stage)) + scale_y_log10() + ggtitle("Asaia")
```

# Mutating the dataframe with padj 0.05 (Significant)
```{r}
res <- res %>% mutate(sig=padj<0.05)
# How many of each?
res %>%  group_by(sig) %>% summarize(n=n())
```

## MA plot with res where padj is < 0.05
```{r}
 res %>% ggplot(aes(baseMean, log2FoldChange, col=sig)) + geom_point() + scale_x_log10() +
  ggtitle("MA plot")
```
 
```{r}
res %>% ggplot(aes(log2FoldChange, -1*log10(pvalue), col=sig)) + geom_point() + ggtitle("Volcano plot")
```

```{r}
resNorm <- lfcShrink(dds, coef=2, type="normal")
resAsh <- lfcShrink(dds, coef=2, type="ashr")
par(mfrow=c(1,3), mar=c(4,4,2,1))
xlim <- c(1,1e5); ylim <- c(-3,3)
plotMA(data.frame(resNorm), xlim=xlim, ylim=ylim, main="normal")
plotMA(data.frame(resAsh), xlim=xlim, ylim=ylim, main="ashr")
```

## R Log transformation rlog(regualrsied log transformation)
```{r}
rld <- rlogTransformation(dds, blind=TRUE)
```

## Distance matrix
```{r}
distsRL <- dist(t(assay(rld)))
mat <- as.matrix(distsRL)
rownames(mat) <-  colData(rld)$sex
colnames(mat) <-  colData(rld)$stage
```

# Coorelation plot
```{r}
hmcol <- colorRampPalette(brewer.pal(9, "Blues"))(255)
heatmap.2(mat, trace="none", col = rev(hmcol),Colv = T,dendrogram = "col")
```

# nbionmWaldTest : BH algorithim
```{r}

DESeq2Table <-  nbinomWaldTest(dds)
DESeq2Res <- results(DESeq2Table, pAdjustMethod = "BH")
table(DESeq2Res$padj < 0.1)
hist(DESeq2Res$pvalue, col = "lavender", main = "True vs False(padj < 0.1)", xlab = "p-values")
```

## MA Plot for above test
```{r}
plotMA(data.frame(DESeq2Res))
```

# RLD / VST / VSD
```{r}
rld <- rlog(dds, blind = FALSE)
#vsd <- vst(dds, blind = FALSE)
dds1 <- estimateSizeFactors(dds)

df <- bind_rows(
  as_data_frame(log2(counts(dds1, normalized=TRUE)[, 1:2]+1)) %>%
    mutate(transformation = "log2(x + 1)"),
  as_data_frame(assay(rld)[, 1:2]) %>% mutate(transformation = "rlog"))
```


```{r}
colnames(df)[1:2] <- c("x", "y")  
 
```

```{r}
ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed() + facet_grid( . ~ transformation) 
```

## Estimate Dispersion
```{r}
countDataSet <- newCountDataSet(data,group)
countDataSet <- estimateSizeFactors(countDataSet)
countDataSet <- estimateDispersions(countDataSet,fitType="local")
plotDispEsts(countDataSet)
```

## nbiom Test
Test between the samples 
```{r}
DEVal_G01 <- nbinomTest(countDataSet,"G0","G1")
DEVal_G02 <- nbinomTest(countDataSet,"G0","G2")
DEVal_G03 <- nbinomTest(countDataSet,"G0","G3")
DEVal_G04 <- nbinomTest(countDataSet,"G0","G4")
DEVal_G12 <- nbinomTest(countDataSet,"G1","G2")
DEVal_G13 <- nbinomTest(countDataSet,"G1","G3")
DEVal_G14 <- nbinomTest(countDataSet,"G1","G4")
DEVal_G23 <- nbinomTest(countDataSet,"G2","G3")
DEVal_G24 <- nbinomTest(countDataSet,"G2","G4")
DEVal_G34 <- nbinomTest(countDataSet,"G3","G4")
```

## Setting up pvale

```{r}
p1 <- subset(DEVal_G01,pval<0.005)
p2 <- subset(DEVal_G02,pval<0.005)
p3 <- subset(DEVal_G03,pval<0.005)
p4 <- subset(DEVal_G04,pval<0.005)
p5 <- subset(DEVal_G12,pval<0.005)
p6 <- subset(DEVal_G13,pval<0.005)
p7 <- subset(DEVal_G14,pval<0.005)
p8 <- subset(DEVal_G23,pval<0.005)
p9 <- subset(DEVal_G24,pval<0.005)
p10 <- subset(DEVal_G34,pval<0.005)
```
## Data binding function

```{r}
dataPlotFunction <- function(binded){
  p_plot <- data.frame(binded[,1])
  nrow(p_plot)
  p_plot <- unique(p_plot)
  colnames(p_plot)<-"taxon_name"
  p <- inner_join(p_plot,dff_merge_all,by="taxon_name")
  rownames(p)<-p$taxon_name
  p <-log(p[,-1])
  return(p)
}
```

## Binding the pvale dataframes
```{r}
binded <- rbind(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10) 
dataPlot <- dataPlotFunction(binded)
nrow(dataPlot)
```

## Heatmap based on binded dataframe (p1-p10)

```{r}
pheatmap(dataPlot[,],fontsize = 5, scale = "none", cluster_rows = TRUE,cutree_rows = 20,cellwidth = 5,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = F, show_colnames = T, main = NA,fontsize_row = 10,ffontsize_col = 8)
```

## plotting by ordering ( Ordering can be by each sample)
```{r}
dataPlot <- dff_merge_all[order(-dff_merge_all$MSP5,-dff_merge_all$MSP6,-dff_merge_all$MSP7,-dff_merge_all$MSP8),]

```

```{r}
pheatmap(log2(dataPlot[5:200,2:20]),fontsize = 5, scale = "none", cluster_rows = TRUE,cutree_rows = 10,cellwidth = 8,
        color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
       cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
        clustering_method = "complete",show_rownames = T, show_colnames = T, main = "5:200 log2" )
```

# Male vs Female study

```{r}
all_male <- dff_merge_all[,c(4,5,8,9,13,16,17,20,21)]
all_female <- dff_merge_all[,c(2,3,6,7,10,11,12,14,15,18,19)]

for_wolbachia_male <- all_male["Asaia",]

wolbachia <- data.frame("Asaia")
names(wolbachia)<-"taxon_name"
for_wolbachia_male <- cbind(wolbachia,for_wolbachia_male)
head(for_wolbachia_male)
Molt <- melt(for_wolbachia_male, id.vars = "taxon_name")
```

## Only Male  (Log2)
```{r}
ggplot(Molt, aes(x = variable, y = log2(value),color=variable,group=taxon_name)) + 
  geom_line() + 
  # facet_wrap(~ Gene_ID)+  
  ggtitle("Wolbachia Read Count ")+
  labs(y="Counts", x = "Sample")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Male-Female Both Asaia & Wolbachia

```{r}
for_wolbachia_male <- all_male[c("Asaia","Wolbachia"),]
for_wolbachia_female <- all_female[c("Asaia","Wolbachia"),]
female_male <- cbind(for_wolbachia_female,for_wolbachia_male)
wolbachia <- data.frame(c("Asaia","Wolbachia"))
names(wolbachia)<-"taxon_name"
for_wolbachia_male <- cbind(wolbachia,female_male)
Molt <- melt(for_wolbachia_male, id.vars = "taxon_name")
ggplot(Molt, aes(x = variable, y = log2(value),color=variable,group=taxon_name)) + 
  geom_line() + 
   facet_wrap(~ taxon_name)+  
  ggtitle("Read Counts ")+
  labs(y="Log2 transformed Counts", x = "Sample : Both")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Between All : Genus Level ( Taking all the samples together)

```{r}
g_data <- read.csv("between_all",sep="\t",header = T)
head(g_data)
rownames(g_data)<-make.unique(as.character(g_data$genus))

pheatmap(log2(g_data[1:50,2:6]),fontsize = 5, scale = "none", cluster_rows = F,cutree_rows = 1,cellwidth = 15,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = T, show_colnames = T, main = "1-50:Log2",fontsize_row = 10,ffontsize_col = 8)

```

## NO G0

```{r}

pheatmap(log2(g_data[10:50,3:6]),fontsize = 5, scale = "none", cluster_rows = F,cutree_rows = 1,cellwidth = 15,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = T, show_colnames = T, main = "10-50:Log2",fontsize_row = 10,ffontsize_col = 8)

pheatmap(log2(g_data[50:100,3:6]),fontsize = 5, scale = "none", cluster_rows = F,cutree_rows = 1,cellwidth = 15,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = T, show_colnames = T, main = "50-100:Log2",fontsize_row = 10,ffontsize_col = 8)

pheatmap(log2(g_data[100:150,3:6]),fontsize = 5, scale = "none", cluster_rows = F,cutree_rows = 1,cellwidth = 15,
         color = colorRampPalette(rev(brewer.pal(n = 10, name ="RdYlBu")))(50),
         cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_distance_cols = "euclidean",
         clustering_method = "complete",show_rownames = T, show_colnames = T, main = "100-150:Log2",fontsize_row = 10,ffontsize_col = 8)
```


```{r}
nrow(g_data)
distsRL <- dist(t(g_data[1:1456,2:6]))
mat <- as.matrix(distsRL)
```

# Coorelation plot with G0
```{r}
hmcol <- colorRampPalette(brewer.pal(9, "Blues"))(255)
heatmap.2(mat, trace="none", col = rev(hmcol),Colv = T,dendrogram = "col")
```

# Coorelation plot without G0

```{r}
head(g_data)
distsRL <- dist(t(g_data[1:1456,3:6]))
mat <- as.matrix(distsRL)
```

# Coorelation plot
```{r}
hmcol <- colorRampPalette(brewer.pal(9, "Blues"))(255)
heatmap.2(mat, trace="none", col = rev(hmcol),Colv = T,dendrogram = "col")
```

## Broad data exploration

<!-- ```{r} -->

<!-- msp_for_plot <- function(df,number){ -->
<!--   by_vs <- df %>% dplyr::summarise(n = n()) -->
<!--   dat_frame <- data.frame(by_vs) -->
<!--   msp_data <- dat_frame[dat_frame$n>number,] -->
<!--   print(nrow(msp_data)) -->
<!--   msp_data <- data.frame(lapply(msp_data,function(x){gsub("MSP","",x)})) -->
<!--   print(nrow(msp_data)) -->
<!--   msp_data$V1 <- as.numeric(as.character(msp_data$V1)) -->
<!--   msp_data <- msp_data[order(msp_data$V1),] -->
<!--   msp_data$V1 <- paste0("MSP",msp_data$V1) -->
<!--   print(nrow(msp_data)) -->
<!--   return(msp_data) -->
<!-- } -->
<!-- mspallya_within <- read.csv("mspallya_within/genus_kaiju_summary.csv",header = F,sep="\t") -->
<!-- mspallya_within <-mspallya_within[,1:11] -->
<!-- ``` -->

### Extracting Function
```{r}
msp_for_plot <- function(df,number){
  by_vs <- df %>% dplyr::summarise(n = n())
  dat_frame <- data.frame(by_vs)
  msp_data <- dat_frame[dat_frame$n>number,]
  print(nrow(msp_data))
  msp_data <- data.frame(lapply(msp_data,function(x){gsub("MSP","",x)}))
  print(nrow(msp_data))
  msp_data$V1 <- as.numeric(as.character(msp_data$V1))
  msp_data <- msp_data[order(msp_data$V1),]
  msp_data$V1 <- paste0("MSP",msp_data$V1)
  print(nrow(msp_data))
  return(msp_data)
}
mspallya_within <- read.csv("mspallya_within/genus_kaiju_summary.csv",header = F,sep="\t")
mspallya_within <-mspallya_within[,1:11]
```

```{r}
head(mspallya_within[,1:11])
```


## V1 vs V6 ( group by Sample and Kingdom 10 as rowSums)

```{r}
group_data <- mspallya_within %>% group_by(V1,V6)#V1,V8  # V1,V11 #V1,V7
head(group_data)
msp_data <- msp_for_plot(group_data,10)
head(msp_data)
nrow(msp_data)
p <- ggplot(msp_data, aes(x = V1, y = n))+
  geom_col(aes(fill = V6),width = 1)+ scale_x_discrete(limit=msp_data$V1) +theme(axis.text.x = element_text(angle = 90))
p
```


## V1 vs V7 (group by sample and phylum with 100 as rowSums)

```{r}
group_data <- mspallya_within  %>% group_by(V1,V7)
msp_data <- msp_for_plot(group_data,100)
head(msp_data)
p <- ggplot(msp_data, aes(x = V1, y = n))+
  geom_col(aes(fill = V7),width = 1)+ scale_x_discrete(limit=msp_data$V1) +theme(axis.text.x = element_text(angle = 90))
p
```

## V1 vs V6 ( group by Sample and Class 100 as rowSums)

```{r}
group_data <- mspallya_within %>% group_by(V1,V8)
head(group_data)
nrow(group_data)
msp_data2 <- msp_for_plot(group_data,100)
head(msp_data2)
nrow(msp_data2)
p <- ggplot(msp_data2, aes(x = V1, y = n))+
  geom_col(aes(fill = V8),width = 3)+ scale_x_discrete(limit=msp_data2$V1) +theme(axis.text.x = element_text(angle = 90))
p
```




```{r}
sessioninfo()
```


